---
title: "ch11.rmd"
author: "Brea Koenes"
date: "10/24/2022"
output: html_document
---

```{r setup, include=FALSE}
library(bayesrules)
library(rstanarm)
library(bayesplot)
library(tidyverse)
library(broom.mixed)
library(tidybayes)
```

# Chapter 11

## 2

a.  

There is no indicator for the Ford category because it is the intercept.

b.  

A car's miles per gallon in a city given its a Subaru

c.  

a Ford's miles per gallon in a city

## 10

a.  

```{r}
penguin_data <- penguins_bayes |> 
  filter(species %in% c("Adelie", "Gentoo"))

summary(penguin_data)
```

b.  

```{r}
require(rstanarm)

model <- stan_glm(
  body_mass_g ~ species + flipper_length_mm, data = penguin_data, 
  family = gaussian,
  prior_PD = TRUE,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 5, iter = 8000, seed = 84735)
```

c.  

```{r}
summary(model)
```

```{r}
# Trace plots of parallel chains
mcmc_trace(model, size = .1)
```

```{r}
# Density plots of parallel chains
mcmc_dens_overlay(model)
```

d.  

```{r}
tidy(model, effects = c("fixed", "aux"),
     conf.int = TRUE, conf.level = 0.95)
```

e.  

```{r}
model2 <- update(model, prior_PD = FALSE)
```

```{r}
# Simulate a set of predictions
set.seed(84735)
model2_predict <- posterior_predict(
  model2,
  newdata = data.frame(species = c("Adelie", "Gentoo"), 
                       flipper_length_mm = 197))

# Plot the posterior predictive models
mcmc_areas(model2_predict) +
  ggplot2::scale_y_discrete(labels = c("Adelie", "Gentoo")) + 
  xlab("flipper_length_mm")
```

## 11

a.  

```{r}
interaction_model <- stan_glm(
  body_mass_g ~ species + flipper_length_mm + species:flipper_length_mm, data = penguin_data, 
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4, iter = 10000*4, seed = 84735)
```

b.  

```{r}
# Posterior summary statistics
tidy(interaction_model, effects = c("fixed", "aux"))
```

```{r}
set.seed(84735)
penguin_data |> na.omit() |>
  add_fitted_draws(interaction_model, n = 50) |>
  ggplot(aes(x = "Adelie", y = "Gentoo", color = flipper_length_mm)) +
    geom_line(aes(y = .value, group = paste(species, .draw)))
```

c.  

```{r}
tidy(interaction_model, effects = c("fixed", "aux"))
```

The interaction terms are necessary for this model. Looking at the summary above, speciesGentoo including the flipper_length_mm interaction has a much smaller standard error than speciesGentoo where the interaction is not factored in.

## 12

a.  

```{r}
three_predictor_model <- stan_glm(
  body_mass_g ~ flipper_length_mm + bill_depth_mm + bill_length_mm, data = penguin_data, 
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4, iter = 10000*4, seed = 84735)
```

b.  

```{r}
posterior_interval(three_predictor_model, prob = 0.95)
```

c.  

All predictors have a significant positive association with body mass.

## 13

a.  

```{r}
mod1 <- stan_glm(
  body_mass_g ~ flipper_length_mm, data = penguin_data, 
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 3, iter = 5000*3, seed = 84735)
```

```{r}
mod2 <- stan_glm(
  body_mass_g ~ species, data = penguin_data, 
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 3, iter = 5000*3, seed = 84735)
```

```{r}
mod3 <- stan_glm(
  body_mass_g ~ flipper_length_mm + species, data = penguin_data, 
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 3, iter = 5000*3, seed = 84735)
```

```{r}
mod4 <- stan_glm(
  body_mass_g ~ flipper_length_mm + bill_length_mm + bill_depth_mm,
  data = penguin_data,
  family = gaussian,
  prior_intercept = normal(42.7, 10, autoscale = TRUE),
  prior = normal(4318, 1700, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 3, iter = 5000*3, seed = 84735)
```

b.  

```{r}
pp_check(mod1)
pp_check(mod2)
pp_check(mod3)
pp_check(mod4)
``` 

c.  

```{r}
penguins_complete <- penguin_data |> 
  select(flipper_length_mm, body_mass_g, species, 
         bill_length_mm, bill_depth_mm) |> 
  na.omit() 
```

```{r}
set.seed(84735)
prediction_summary_cv(model = mod1, data = penguins_complete, k = 10)  
```

```{r}
set.seed(84735)
prediction_summary_cv(model = mod2, data = penguins_complete, k = 10)  
```

```{r}
set.seed(84735)
prediction_summary_cv(model = mod3, data = penguins_complete, k = 10)  
```

```{r}
set.seed(84735)
prediction_summary_cv(model = mod4, data = penguins_complete, k = 10)  
```

d.  

```{r}
# Calculate ELPD for the 4 models
set.seed(84735)
loo_1 <- loo(mod1)
loo_2 <- loo(mod2)
loo_3 <- loo(mod3)
loo_4 <- loo(mod4)

# Results
c(loo_1$estimates[1], loo_2$estimates[1], 
  loo_3$estimates[1], loo_4$estimates[1])
```

```{r}
# Compare the ELPD for the 4 models
loo_compare(loo_1, loo_2, loo_3, loo_4)
```

The best model is model 4. mod3 is -40 worse than 4, with model 1 a similar amount lower as well. Model 2 is the worst model comparatively.

e.  

Looking at the ELPD posterior predictive accuracy, 10-fold cross-validation, and pp_check() charts, model 4 is the best model. It has the best accuracy and association between predictors and response variable. This makes sense that this is the best model; it uses more predictors than the other models, and all of the predictors are relevant. 

## 15

```{r}
summary(weather_WU)
```

```{r}
weather_model <- stan_glm(
  temp3pm ~ location + rainfall + pressure9am + month,
  data = weather_WU, family = gaussian,
  prior_intercept = normal(30, 5),
  prior = normal(0, 3, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 4, iter = 5000*2, seed = 84735)
```

```{r}
summary(weather_model)
```

```{r}
# Trace plots of parallel chains
mcmc_trace(weather_model, size = .1)
```

```{r}
# Density plots of parallel chains
mcmc_dens_overlay(weather_model)
```
