---
html_document:
  code_download: yes
author: "Brea Koenes"
date: "10/19/2022"
fig_width: 8
fig_height: 3
output:
  html_document:
    df_print: paged
title: "Stat341 Portfolio"
---

```{r setup, include=FALSE}
library(bayesrules)
library(tidyverse)
library(ggformula)
library(dplyr)
```

# Chapter 4

## 4.5

#### Kimya Beta(1, 2)

```{r}
set.seed(84735)
kimya_sim <- data.frame(pi = rbeta(10000, 1, 2)) |> 
  mutate(y = rbinom(10000, size = 7, prob = pi))

kimya_posterior <- kimya_sim |> 
  filter(y == 3)

ggplot(kimya_posterior, aes(x = pi)) + 
  geom_histogram()
```

```{r}
kimya_posterior |>
  summarize(mean(pi), sd(pi))
```

#### Fernando Beta(0.5, 1)

```{r fernando c}
set.seed(84735)
Fernando_sim <- data.frame(pi = rbeta(10000, 0.5, 1)) |> 
  mutate(y = rbinom(10000, size = 7, prob = pi))

Fernando_posterior <- Fernando_sim |> 
  filter(y == 3)

ggplot(Fernando_posterior, aes(x = pi)) + 
  geom_histogram()
```

```{r}
Fernando_posterior |>
  summarize(mean(pi), sd(pi))
```

#### Ciara Beta(3, 10)

```{r fernando c}
set.seed(84735)
Ciara_sim <- data.frame(pi = rbeta(10000, 3, 10)) |> 
  mutate(y = rbinom(10000, size = 7, prob = pi))

Ciara_posterior <- Ciara_sim |> 
  filter(y == 3)

ggplot(Ciara_posterior, aes(x = pi)) + 
  geom_histogram()
```

```{r}
Ciara_posterior |> 
  summarize(mean(pi), sd(pi))
```

#### Taylor Beta(2, 0.1)

```{r}
set.seed(84735)
Taylor_sim <- data.frame(pi = rbeta(10000, 1, 2)) |> 
  mutate(y = rbinom(10000, size = 7, prob = pi))

Taylor_posterior <- Taylor_sim |> 
  filter(y == 3)

ggplot(Taylor_posterior, aes(x = pi)) + 
  geom_histogram()
```

```{r}
Taylor_posterior |>
  summarize(mean(pi), sd(pi))
```

## 4.14

a.  The posterior mode of pi as the weighted average of the prior mode and observed sample success rate:

$$ \begin{align*}

Mode(\pi∣Y=y) &= \frac{\alpha+y−1}{\alpha+\beta+n−2} \\
&= \frac{\alpha−1}{\alpha+\beta+n−2} + \frac{y}{\alpha+\beta+n−2} \\
&= \frac{\alpha−1}{\alpha+\beta+n−2} ⋅ \frac{\alpha+\beta−2}{\alpha+\beta−2} + \frac{y}{\alpha+\beta+n−2} ⋅ \frac{n}{n} \\
&= \frac{\alpha+\beta−2}{\alpha+\beta+n−2} ⋅ \frac{\alpha−1}{\alpha+\beta−2} + \frac{n}{\alpha+\beta+n−2} ⋅ \frac{y}{n} \\
&= \frac{\alpha+\beta−2}{\alpha+\beta+n−2} ⋅ Mode(\pi) + \frac{n}{\alpha+\beta+n−2} ⋅ \frac{y}{n} \\

\end{align*} $$

b.

Using the strong law of large numbers and given that there are appropriate regularity conditions, the posterior mode converges to 1 as n increases.

# Chapter 5

## 5.10

a.  

```{r}
data <- c(-0.7,1.2,4.5,-4)
plot_normal_likelihood(y = data, sigma = 2)
```

b.  

```{r}
plot_normal_normal(mean = 7, sd = 2.6, sigma = 2, y_bar = .25, n = 4)
```

c.  

```{r}
bayesrules::summarize_beta(alpha = 2, beta = 7)
```

```{r}
summarize_normal_normal(mean = 7, sd = 2.6, sigma = 2, y_bar = .25, n = 4)
```

d.  

My understanding of the mean evolved from the prior by the posterior's mean being below .25 while the mean of the prior was .25.

e.  

```{r}
pnorm(0, mean=7,sd=2, log.p=FALSE)
```

f.  

```{r}
pnorm(8, mean=7,sd=2, log.p=FALSE)
```

## 5.11

a.  

```{r}
summarize_normal_normal(mean = 80, sd = 4, sigma = 3, y_bar = 86, n = 32)
```

b.  

```{r}
summarize_normal_normal(mean = 80, sd = 4, sigma = 3, y_bar = 82, n = 32)
```

c.  

```{r}
summarize_normal_normal(mean = 80, sd = 4, sigma = 3, y_bar = 84, n = 32)
```

## 5.13

a.  

```{r}
plot_normal(mean = 30, sd = 5)
```

b.  

```{r}
ggplot(weather_perth, aes(x = temp3pm)) + 
  geom_density()
```

It is reasonable to assume a Normal model for the temperature data. We can assume this because there is a symetrical bell-shaped distribution.

c.  

```{r}
weather_perth |>
  summarize(mean(temp3pm))
```

```{r}
summarize_normal_normal(mean = 30, sd = 5, sigma = 3, 
                        y_bar = 24, n = 1000)
```

d.  

```{r}
plot_normal_normal(mean = 30, sd = 5, sigma = 3,
                   y_bar = 24, n = 1000)
```

Given the prior, I originally thought that the average temperature is around 30 degrees Celsius. However, there was uncertainty in that prior since it could be anywhere from 10 to 50 degrees. Due to this uncertainty, I was not surprised to see that the posterior is centered around 24. 
