---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
# Load packages
library(bayesrules)
library(tidyverse)
library(ggformula)
library(dplyr)
```

# Chapter 9

## 9

a.

```{r}
summarize_normal_normal(mean = 1000, sd = 2000, sigma = 10,
                        y_bar = 9000, n = 10000)
```

b.

```{r}
require(rstanarm)

model <- stan_glm(
  rides ~ humidity, data = bikes, 
  family = gaussian,
  prior_PD = TRUE,
  prior_intercept = normal(1000, 2000, autoscale = TRUE),
  prior = normal(0, 10, autoscale = TRUE), 
  prior_aux = exponential(1, autoscale = TRUE),
  chains = 5, iter = 8000, seed = 84735)
```

c.

```{r}
require(bayesplot)

# Trace plots of parallel chains
mcmc_trace(model, size = 100)

# Density plots of parallel chains
mcmc_dens_overlay(model)
```

d.

Ridership decreases as humidity increases. There is much variability with the ridership per day, so that impacts the strength of the relationship.

## 10

a.

```{r}
# Load and plot data
data(bikes)
ggplot(bikes, aes(x = humidity, y = rides)) + 
  geom_point(size = 0.5) + 
  geom_smooth(method = "lm", se = FALSE)
```

b.

The data is semi-normally distributed. It's not evenly distributed across the x-axis, but the points are evenly distributed across the y-axis. Since it doesn't form a trumpet-like shape, it could be fine for Normal regression. 

## 11

a.

b.

c.

## 12

a.

b.

c.

d.

## 13

a.

b.

c.

d.

## 16

a.

b.

c.

d.

## 17

a.

b.

## 18

a.

b.

c.

d.

e.

## 19

a.

b.

c.

d.

e.

# Chapter 10

## 7 

a.

b.

## 13

a.

b.

## 14

a.

b.

c.

d.

e.

## 15

a.

b.

c.

d.

## 16

a.

b.

c.

d.

## 17

a.

b.

c.

## 18

## 19

a.

b.

c.

d.

